Canonical rings of rational divisors on projective space and Hirzebruch surfaces

### 2015-07-15 surfaces-write-up/surface-write-up.tex
added intro over my pay-grade/head

\section{Introduction}
\todo{Clean up this stuff on siegel modular forms, hilbert modular forms, say something about hasse keel program, canoncical rings studied before}

\todo{Might be reasonable to think about higher dimensional rational varieties that are similar to these hirzebruch surfaces.}

\todo{The following stuff is almost nonsensical now, will be edited much more soon}

Hasse keel program, birational map from $proj R(K + \Delta)$ to $M_2$, they showed finite generation, we're giving explicit generators and bounds.

The work in this paper also yields bounds for certain rings of Hilbert modular forms. Recall that a Hilbert modular form is a generalization of modular forms to multiple variables.

Siegel modular forms are a generalization of modular forms on the Siegel upper half plane to automorphic forms on the Siegel upper half space. Explicitly, the Siegel uper half space is $\mathscr H \cong \bc^n/\Lambda$ and siegel modular forms are functions on $Sp_{2g} \backslash \mathscr H.$ Then, modular functions are differentials invariant under the action of $Sp_{2g}$. In the case that $\Gamma \backslash \mathscr H$.

Example: the Siegel 2-upper half plane $\bc^2/\Lambda$ Note that $\bc^2/\Lambda$ is algebraic if and only if there exists a hermitial form on $\Lambda$. Then, $Sp_{2g}$ acts on algebraic lattices. The quotient by this gives you the space of abelian varieties. This might be an open subset of projective space, according to David, 65$\%$.

@@

\noindent
Suppose $e_1, \ldots, e_t \in \Sigma$ with $e_i = (\pdeg_i, c_1^i,
\ldots, c_n^i)$ are a set of {\bf extremal rays} of $\Sigma$,
meaning that the $\Sigma$ is contained in the $R_{\geq 0}$ span of
$e_1, \ldots, e_t$.  Then, as a semigroup, $\Sigma$ is generated by
elements whose first coordinate is less than $\sum_{i = 1}^{n} \pdeg_i$
. Furthermore, every element $\sigma \in \Sigma$ can be written in
a canonical form $\lambda + \sum_{i=1}^{t} w_i e_i,$ with the first
coordinate of $\lambda$ less than $\sum_{i=1}^{t}\pdeg_i$ and $w_1,
\ldots, w_t \in \bz$.
\end{prop}

@@

\begin{lem}
\label{lem:bound-ker-chi}
Retaining the notation of Lemma ~\ref{lem:composite-map}, if $\Sigma$ has extremal rays $e_1,\ldots, e_r$ in degrees $d_1, \ldots, d_r$ then $\ker \chi$ has a system of generators in degrees at most $2(\sum_{i=1}^{r}d_r-1)$.
\end{lem}
\begin{proof}
Since $e_1, \ldots, e_r$ are extremal rays, by Proposition ~\ref{prop:cone-generation}, every element $\sigma \in \Sigma$ can be written in a canonical form $\lambda + \sum_{i=1}^{r}e_i$. Now, letting $\lambda_0 := 0,\lambda_1, \ldots, \lambda_m$ be all elements of $\Sigma$ so that we can write $\lambda_i = \sum_{i=1}^{r}\bida_i e_i$ with $0 \leq \bida_i < 1$ for all $i$. Then, for any $1 \leq i \leq j \leq m,$ we can write $\lambda_i + \lambda_j$ in the above canonical form, yielding a relation in degree at most $\deg \lambda_i + \deg \lambda_j \leq 2 \cdot \left( \sum_{i=1}^{r}d_r -1 \right).$ Furthermore, these relations generate all relations, as one can apply a sequence of these relations to put any $\sigma \in \Sigma$ into canonical form.
\end{proof}

@@

Write
\begin{align*}
	D = \sum_{i=0}^{n}\alpha_i D_i.
\end{align*}
where $\alpha_i \in \bq$ and $\deg D_i = \bida_i$. Let $f_i$ be Cartier divisors such that $D_i = V(f_i)$. \todo{Should this say $D_i$ is a Cartier divisor.  Also should we say $\deg(f_i)= \bida_i$ rather than $\deg(D_i)$.}
We shall further assume that (after reordering) $f_0,\ldots, f_{m}$ form a basis for degree one rational functions on $\bp^m$. \todo{Peter: Make reference to ghost points here?}

\begin{prop}
\label{prop:pm-span-and-basis}
The functions $u^d \cdot \prod_{i=1}^n f_i^{c_i}$ satisfying
\begin{align}
\label{align:pm-span}
\sum_{i=1}^{n} c_i \cdot \bida_i \cdot f_i = 0 && \text{ and } &&c_i \geq \lfloor \alpha_i d\rfloor	
\end{align}

@@

We first restrict attention to the case $n=1$, so $D = \alpha D_i$.

@@

\begin{rem}
When $\langle f_1, \ldots, f_r \rangle = \langle x_0, \ldots, x_n \rangle$ (the irrelevant ideal), we may directly use the hypersurface method by considering $x_0, \ldots, x_n$ instead of the $f_i$'s. However, when $\sqrt{\langle f_1, \ldots, f_r \rangle} = \langle x_0, \ldots, x_n \rangle$ (or the saturation), we cannot obtain the same result from our modification of O'Dorney's method as seen in Example ~\ref{eg:radical}. Instead we must use ghost points and hypersurfaces.
\end{rem} \todo{Peter: What is the purpose of this remark?  The only method we have extends to ghost points so that this is satisfied.}

@@

This example further generalizes to the following situation: Suppose $D = \sum_{i=1}^{n} \frac{\num_i}{\den_i}D_i \in \di \bp^m \otimes \bq,$ where $\deg D_i = \bida_i$ and \todo{Peter: We need some condition that $D_i$'s be irreducible divisors or something.} $\deg D = \frac{1}{\lcm_{1 \leq i \leq n}(\den_i \cdot \bida_i)}$. Then, if $D_i = V(f_i)$ where all $f_i$ can be written as a polynomial function in $x_0, \ldots, x_{m-1},$ we have $R_D$ always has a generator in degree $\lcm(\den_i \cdot \bida_i)$.
\end{example}

\begin{rem}
\label{rem:ghost-motivation}
Example ~\ref{eg:hyperplane} shows that the naive generalization of ~\cite[Theorem 8]{dorney:canonical} of generation in degree at most $\sum_{i=1}^{n}\ell_i$ cannot possibly hold. The reason for this is that the divisors may ``be contained in a hyperplane,'' i.e., be expressable as functions in $m$ of the $m+1$ variables on $\bp^m$. This problem can easily be circumvented by adding in ``ghost points.'' That is, we may add divisors of the form $0 \cdot H_i$ to $D$, and reorder so that if $D = \sum_{i=1}^{n}\alpha_i V(f_i)$, then $f_0, \ldots, f_n$
\end{rem}

@@

We now move on offer bounds on generators and relations of $R_D$ for a general divisor $D$ on $\bp^m$.

Let $D = \sum_{i=1}^n \alpha_i C_i$ for curves $C_i$ on $\mathbb{P}^m$ and let $f_i$ be homogeneous polynomials in $x_0, \ldots, x_m$ such that $C_i = v(f_i)$ for all $i$; suppose $\deg(f_i) = a_i$. 
\begin{lem} \label{P-m-generators}
$R_D$ is generated in degree at most $\sum_{i=1}^n \ell_i a_i$.
% and relations in degree at most  .\todo{insert numbers}
\end{lem}
\begin{proof}
Let 
\begin{align}\label{eqn:Sigma-def}
	\Sigma = \{(d, c_1, \ldots, c_n) \in \bz^{n+1} : c_i \geq - d \alpha_i, \; 1 \leq i \leq n, \text{ and } \sum_{i=1}^{n}k_i^j = 0, 1 \leq j \leq t\}.
\end{align}
We note that $\Sigma$ has extremal rays given by the lattice points 
\begin{equation}\label{defn:e_i-P-n}
	e_i = (\ell_i \bida_i, - \alpha_1 \ell_i \bida_i, \ldots -\alpha_{i-1} \ell_i \bida_i, \ell_i \sum_{j\ne i} \alpha_j \bida_j, -\alpha_{i+1} \ell_i \bida_i, \ldots, -\alpha_n, \ell_i \bida_i)
\end{equation}
for each $i\in \{1, \ldots n\}$.
Viewed inside of $R_D$, we have $e_i$ corresponding to
\begin{equation}\label{eqn:epilon-def-P-m}
	\epsilon_i = \prod_{j\ne i} (f_j)^{-\alpha_j \ell_i a_i} {f_i}^{\ell_i \sum_{j\ne i} \alpha_j a_j}\in R_D.
\end{equation}
Therefore, applying proposition \ref{prop:cone-generation}, we see $R_D$ is generated in degrees less than
\[
	\sum_{i=1}^n \ell_i a_i.
\]
\end{proof}

Let $x_1, \ldots x_r$ be the generators in degrees at most $\sum_{i=1}^n \ell_i \bida_i$ (given by lemma \ref{P-m-generators}), and let $\phi$ be the surjection $k[x_1, \ldots x_r] \to R_D$.  We can factor $\phi$ through 
\[
	\Sigma =  \langle u^d z_1^{c_1} \cdots z_n^{c_n} : c_i \in \mathbb{Z}, \; c_i \geq -d \alpha_i, \mbox{ and }\sum_{i=1}^{n} \bida_i c_i = \sum_{i=1}^{n} \bidb_i c_i. \rangle 
\]
by
\[
\begin{tikzcd}
\bk[x_1,\ldots, x_r] \ar {r}{\chi} & \bk[\Sigma] \ar {r}{\psi} & R_D \\
x_i \ar r & u^{d_i}z_1^{c_{i1}} \cdots z_n^{c_{in}} \ar r & u^{d_i}f_1^{c_{i1}} \cdots f_n^{c_{in}}.
\end{tikzcd}
\]

By Lemma \ref{lem:bound-ker-chi}, we can bound the degree of generation of $\ker(\chi)$ below
\[
	2 \sum_{i=1}^n \ell_i a_i.
\]
Finally, we calculate the degree of generation of $\psi$:

\begin{lem}\label{P-m-relations-psi}
$\ker(\psi)$ is generated in degree less than
\[
	\frac{\max_{1\le i \le n}(\bida_i)}{\deg(D)} +  \sum_{i=1}^n \ell_i a_i.
\]
\end{lem}
\begin{proof}
We first claim that $\psi$ there are $\beta_1, \ldots, \beta_n$ such that $\psi$ has relations of the form
\begin{equation}\label{eqn:relations-psi}
	(z_i - \beta_i)\prod_{j=1}^n {z_j}^{c_{j}}
\end{equation}

\noindent
for all $d\in \mathbb{N}$ and $c_i \ge -\alpha_i d$ satisfying $\bida_j + \sum_{i = 1}
^n \bida_i c_i = 0$.

To find such a $\beta$, we simply write $\epsilon_j$ as a 
polynomial in $\epsilon_1, \ldots \epsilon _3$. Furthermore, these
generate all relations, since they allow us to reduce any $\prod_{i =
1}^n \epsilon_i^{r_i}$ to a canonical form as given by Proposition \ref{prop:cone-generation}.

We now fix $z_i$ and bound the degree of generation of this set of relations of the form $(z_i - \beta_i) \prod_{j=1}^n z_j^{c_j}$.  We first associate $(z_i - \beta_i)\prod_{j=1}^n z_j^{c_j}$ with $(d, c_1, \ldots, c_n)$ where $d$ is its degree.  Let $\Sigma_i \subseteq \mathbb{Z}^{n+1}$ be the set of points of the form $(d, r_1, \ldots, r_n)$ satisfying $r_j \ge -d \alpha_j$ for all $j$ and $\sum_{j=1}^n r_j a_j = -a_i$.  
Then we see $\Sigma_i$ is contained in the $\mathbb{R}_{\ge 0}$-span of $e_1, \ldots, e_n$ (as defined in Equation \ref{defn:e_i-P-n}) translated by by $\delta_i = (\frac{a_i}{\deg(D)}, -\frac{\alpha_1 a_1}{\deg(D)}, \ldots, - \frac{\alpha_n a_n}{\deg(D)})$.

Therefore, we can write any element of $\Sigma'$ as
\[
	\delta_i + \sum_{j=1}^n r_j e_j
\]
where $r_j \in \mathbb{R}$ for each $j$.

Whenever some $r_j \ge 1$, we can generate the relation at hand by one of lower degree by factoring out a copy of $\epsilon_j$ (see Equation \ref{eqn:epilon-def-P-m}).  Therefore, $\ker(\psi)$ is generated in degrees less than
\[
	\frac{\bida_i}{\deg(D)} + \sum_{i=1}^n \ell_i a_i.
\]
\end{proof}

Therefore, since $\ker(\chi)$ and $\ker(\psi)$ are generated in degrees less than $2\sum_{i=1}^n \ell_i a_i$ and $\delta_i + \sum_{i=1}^n \ell_i a_i$ respectively, Lemma \ref{lem:composite-map} implies that $\ker(\phi)$ is generated in degrees less than
\[
	\max \left(2 \sum_{i=1}^n \ell_i a_i, \frac{\max_{1\le i \le n}(\bida_i)}{\deg(D)} + \sum_{i=1}^n \ell_i a_i \right).
\]

@@

\begin{thm}\label{P-m-generators/relations}
Let $D = \sum_{i=1}^n \alpha_i C_i$ for curves $C_i$ on $\mathbb{P}^m$.  Let $f_i$ be homogeneous polynomials in $x_0, \ldots, x_m$ such that $C_i = v(f_i)$ for all $i$; suppose $\deg(f_i) = a_i$. 
Then $R_D$ is generated in degree at most 
\[
	\sum_{i=1}^n \ell_i a_i
\]
with relations generated in degree at most
\[
	\max \left(2 \sum_{i=1}^n \ell_i a_i, \frac{\max_{1\le i \le n}(\bida_i)}{\deg(D)} + \sum_{i=1}^n \ell_i a_i \right).
\]
\end{thm}

\section{Canonical Rings on Hirzebruch surfaces}
Let $D=\sum_{i=1}^n \alpha_i C_i \in \mathbb{Q} \otimes \di(\hirz_
k)$ for some irreducible curves $C_1, \ldots C_n$ of $\hirz_k$, such
that $C_1, \ldots, C_4$ are independent curves with bi-degrees $(1,0)
, (1,0), (0,1)$, and $(0,1)$ respectively. Let $C_i = V(f_i)$ where
$f_i \in \mathscr{O}(\bida_i, \bidb_i)$.  Suppose $f_1$ and $f_2$ are independent linear
polynomials in $x_0$ and $x_1$ and $f_3$ and $f_4$ are independent
linear polynomials in $y_0$ and $y_1$.

\begin{defn}
Define 
\begin{equation}\label{eqn:define-T=(D)}
	\Te(D) = \left\{i \in \{1, \ldots, n\}: \bida_i \sum_{k=1}^n \bidb_k 
\alpha_k = \bidb_i \sum_{k=1}^n \bida_k \alpha_k \right\},
\end{equation}

\begin{equation}\label{eqn:define-T+(D)}
	\Tp(D) = \left\{ i \in \{1, \ldots, n\}:  \vphantom{\sum_{k=1}^n} 
\bida_i \sum_{k=1}^m \alpha_k \bidb_k > \bidb_i \sum_{k=1}^n \alpha_k \bida_k 
\right\},
\end{equation}

\noindent
and
\begin{equation}\label{eqn:define-T-(D)}
	\Tm(D) = \left\{ j \in \{1, \ldots, n\}: \bida_j \sum_{k=1}^n \alpha_
k \bidb_k < \bidb_j \sum_{k=1}^n \alpha_k \bida_k s \right\}.
\end{equation}
\end{defn}

>>>>

\section{Introduction}

\todo{write more about hassett keel program}
\cite{hassett:classical-and-minimal-models}
Hasse keel program, birational map from $proj R(K + \Delta)$ to $M_2$, they showed finite generation, we're giving explicit generators and bounds.

The work in this paper also yields bounds for certain rings of Hilbert 
modular forms and Siegel modular forms. Recall that Hilbert and Siegel 
modular forms are both to higher dimensions variables. In the case of 
Hilbert Modular forms, one considers modular forms on an $m$ fold product 
of upper half planes, which transform under the action of finite index 
subgroups of the groups of positive determinant matrices. Similarly, Siegel 
modular forms are modular forms on the Siegel upper half space which are 
stable under the action of the symplectic group. While in general these 
quotients can be quite complicated spaces, in the case that the resulting 
quotient spaces are isomorphic to $\bp^m$ for some $m$, or some Hirzebruch 
surface, our main result gives a method for bounding the degree of 
generators and relations for the corresponding rings of Hilbert modular 
forms and Siegel modular forms. Namely, as in \cite{vzb:stacky}, one can 
port these modular form groups from the analytic category to the algebraic 
category, and then view the ring of modular forms as a certain stacky 
canonical ring on our underlying algebraic space.
\todo{include references}

\todo{Possibly keep or delete following example}
Example: the Siegel 2-upper half plane $\bc^2/\Lambda$ Note that $\bc^2/\Lambda$ is algebraic if and only if there exists a hermitial form on $\Lambda$. Then, $Sp_{2g}$ acts on algebraic lattices. The quotient by this gives you the space of abelian varieties. This might be an open subset of projective space, according to David, 65$\%$.

@@

\noindent
Let $e_1, \ldots, e_t \in \Sigma,$ with $e_i = (\pdeg_i, c_1^i,
\ldots, c_n^i)$ be a set of {\bf extremal rays} of $\Sigma$,
meaning that the $\Sigma$ is contained in the $R_{\geq 0}$ span of
$e_1, \ldots, e_t$.  Then, as a semigroup, $\Sigma$ is generated by
elements whose first coordinate is less than $\sum_{i = 1}^{n} \pdeg_i$
. Furthermore, every element $\sigma \in \Sigma$ can be written in
a canonical form $\lambda + \sum_{i=1}^{t} w_i e_i,$ with the first
coordinate of $\lambda$ less than $\sum_{i=1}^{t}\pdeg_i$ and $w_1,
\ldots, w_t \in \bz$.
\end{prop}

@@

\begin{lem}
\label{lem:bound-ker-chi}
Retaining the notation of Lemma ~\ref{lem:composite-map}, if $\Sigma$ has extremal rays $e_1,\ldots, e_r$ in degrees $d_1, \ldots, d_r$ then $\ker \chi$ has a system of generators in degrees less than $2(\sum_{i=1}^{r}d_r-1)$.
\end{lem}
\begin{proof}
Since $e_1, \ldots, e_r$ are extremal rays, by Proposition ~\ref{prop:cone-generation}, every element $\sigma \in \Sigma$ can be written in a canonical form $\lambda + \sum_{i=1}^{r}e_i$. Now, letting $\lambda_0 := 0,\lambda_1, \ldots, \lambda_m$ be all elements of $\Sigma$ so that we can write $\lambda_i = \sum_{i=1}^{r}\bida_i e_i$ with $0 \leq \bida_i < 1$ for all $i$. Then, for any $1 \leq i \leq j \leq m,$ we can write $\lambda_i + \lambda_j$ in the above canonical form, yielding a relation in degree at most $\deg \lambda_i + \deg \lambda_j \leq 2 \cdot \left( \sum_{i=1}^{r}d_r -1 \right).$ Furthermore, these relations generate all relations, as one can apply a sequence of these relations to put any $\sigma \in \Sigma$ into canonical form.
\end{proof}

@@

Write
\begin{align*}
	D = \sum_{i=0}^{n}\alpha_i D_i.
\end{align*}
where $\alpha_i \in \bq$ and $\deg D_i = \bida_i$. Let $f_i$ be Cartier divisors such that $D_i = V(f_i)$. We shall further assume that (after reordering) $f_0,\ldots, f_{m}$ form a basis for degree one rational functions on $\bp^m$. \todo{Peter: Make reference to ghost points here?}
\todo{Peter: Also, this is a poorly placed statement of our assumption, because we are not assuming that for the one point/induction case that follows right?}

\begin{prop}
\label{prop:pm-span-and-basis}
The functions $u^d \cdot \prod_{i=1}^n f_i^{c_i}$ satisfying
\begin{align}
\label{align:pm-span}
\sum_{i=0}^{n} c_i \cdot \bida_i \cdot f_i = 0 && \text{ and } &&c_i \geq \lfloor \alpha_i d\rfloor	
\end{align}

@@

\todo{the following bound is the generation bound beginning from evan's paper}

@@

\begin{rem}
When $\langle f_1, \ldots, f_r \rangle = \langle x_0, \ldots, x_n \rangle$ (the irrelevant ideal), we may directly use the hypersurface method by considering $x_0, \ldots, x_n$ instead of the $f_i$'s. However, when $\sqrt{\langle f_1, \ldots, f_r \rangle} = \langle x_0, \ldots, x_n \rangle$ (or the saturation), we cannot obtain the same result from our modification of O'Dorney's method as seen in Example ~\ref{eg:radical}. Instead we must use ghost points and hypersurfaces.
\end{rem}

@@

This example further generalizes to the following situation: Suppose $D = \sum_{i=1}^{n} \frac{\num_i}{\den_i}D_i \in \di \bp^m \otimes \bq,$ where $\deg D_i = \bida_i$ and $\deg D = \frac{1}{\lcm_{1 \leq i \leq n}(\den_i \cdot \bida_i)}$. Then, if $D_i = V(f_i)$ where all $f_i$ can be written as a polynomial function in $x_0, \ldots, x_{m-1},$ we have $R_D$ always has a generator in degree $\lcm(\den_i \cdot \bida_i)$.
\end{example}

\begin{rem}
\label{rem:ghost-motivation}
Example ~\ref{eg:hyperplane} shows that the naive generalization of ~\ref[Theorem 8]{dorney:canonical} of generation in degree at most $\sum_{i=1}^{n}\ell_i$ cannot possibly hold. The reason for this is that the divisors may ``be contained in a hyperplane,'' i.e., be expressable as functions in $m$ of the $m+1$ variables on $\bp^m$. This problem can easily be circumvented by adding in ``ghost points.'' That is, we may add divisors of the form $0 \cdot H_i$ to $D$, and reorder so that if $D = \sum_{i=1}^{n}\alpha_i V(f_i)$, then $f_0, \ldots, f_n$
\end{rem}

@@

Let $D = \sum_{i=1}^n \alpha_i C_i$ for curves $C_i$ on $\mathbb{P}^m$.  Let $f_i$ be homogeneous polynomials in $x_0, \ldots, x_m$ such that $C_i = v(f_i)$ for all $i$; suppose $\deg(f_i) = a_i$. 
\begin{lem} \label{P-m-generators}
$R_D$ is generated in degree at most $\sum_{i=1}^n \ell_i a_i$.
% and relations in degree at most  .\todo{insert numbers}
\end{lem}
\begin{proof}
Let 
\begin{align}\label{eqn:Sigma-def}
	\Sigma = \{(d, c_1, \ldots, c_n) \in \bz^{n+1} : c_i \geq - d \alpha_i,1 \leq i \leq n \text{ and } \sum_{i=1}^{n}k_i^j = 0, 1 \leq j \leq t\}.
\end{align}
We note that $\Sigma$ has extremal rays given by the lattice points $e_i$ corresponding to
\begin{equation}\label{eqn:epilon-def-P-m}
	\epsilon_i = \prod_{j\ne i} (f_j)^{-\alpha_j \ell_i a_i} {f_i}^{\ell_i \sum_{j\ne i} \alpha_j a_j}
\end{equation}
for each $i\in \{1, \ldots n\}$.  Therefore, applying proposition \ref{prop:cone-generation}, we see $R_D$ is generated in degree less than
\[
	\sum_{i=1}^n \ell_i a_i.
\]
\end{proof}

Let $x_1, \ldots x_r$ be the generators given by lemma  \ref{P-m-generators}, and $\phi$ be the surjection $k[x_1, \ldots x_r] \to R_D$.  We can factor $\phi$ through 
\[
	\Sigma =  \langle u^d z_1^{c_1} \cdots z_n^{c_n} : c_i \geq -d \alpha_i, \sum_{i=1}^{n} \bida_i c_i = \sum_{i=1}^{n} \bidb_i c_i. \rangle 
\]
into
\[
\begin{tikzcd}
\bk[x_1,\ldots, x_r] \ar {r}{\chi} & \bk[\Sigma] \ar {r}{\psi} & R_D \\
x_i \ar r & u^{d_i}z_1^{c_{i1}} \cdots z_n^{c_{in}} \ar r & u^{d_i}f_1^{c_{i1}} \cdots f_n^{c_{in}}.
\end{tikzcd}
\]

Applying Lemma \ref{lem:bound-ker-chi}, we can bound the degree of generation of $\ker(\chi)$ less than
\[
	\sum_{i=1}^n \ell_i a_i.
\]
Finally, we calculate the degree of generation of $\psi$:

\begin{lem}\label{P-m-relations-psi}
$\ker(\psi)$ is generated in degree at most 
\[
	\delta_i + \sum_{i=1}^n \ell_i a_i.
\]
\end{lem}
\begin{proof}
We first claim that $\psi$ has a relation of the form
\begin{equation}\label{eqn:relations-psi}
	(z_i - \beta_j)\prod_{j=1}^n {z_j}^{c_{j}}
\end{equation}

\noindent
for some $d\in \mathbb{N}$, $c_i \ge -\alpha_i d$, $\bida_j + \sum_{i = 1}
^n \bida_i c_i = 0$, $\bidb_j + \sum_{i=1}^n \bidb_i c_i = 0$, and $\beta_j \in
k[z_1, \ldots, y_3]$ satisfies $\psi(\beta_j)\in \mathscr{O}(\bida)$.

To find such a $\beta$, we simply write $\epsilon_j$ as a 
polynomial in $\epsilon_1, \ldots \epsilon _3$. Furthermore, these
generate all relations, since they allow us to reduce any $\prod_{i =
1}^n \epsilon_i^{r_i}$ to a canonical form. \todo{add reference to
earlier description of canonical form.}

Now we bound the degree of generation of these relations for a fixed $z_i$.  We associate $(z_i - \beta_j)\prod_{j=1}^n z_j^{c_j}$ that has degree $d$ with the lattice point $(d, c_1, \ldots, c_n)$.  Let $\Sigma_i \subseteq \mathbb{R}^{n+1}$ be the set of points of the form $(d, r_1, \ldots, r_n)$ satisfying $r_j \ge -d \alpha_j$ for all $j$ and $\sum_{j=1}^n r_j a_j = -a_i$.  Then we see $\Sigma_i$ is a translation of $\Sigma$ (see Equation \ref{eqn:Sigma-def}) by $\delta_i = (\frac{a_i}{\deg(D)}, -\frac{\alpha_1 a_1}{\deg(D)}, \ldots, - \frac{\alpha_n a_n}{\deg(D)})$.

Therefore, we can write any element of $\Sigma'$ as
\[
	\delta_i + \sum_{j=1}^n r_j e_j
\]
for $r_j \in \mathbb{R}$ for each $j$.

Whenever some $r_j \ge 1$, we can generate the generation at hand by one of lower degree by factoring out the section $\epsilon_j$ corresponding to the lattice point $e_j$ (see Equation \ref{eqn:epilon-def-P-m}).  Therefore, we can bound the degree of generation of $\ker(\psi)$ less than
\[
	\delta_i + \sum_{i=1}^n \ell_i a_i.
\]
\end{proof}

Therefore, since $\ker(\chi)$ and $\ker(\psi)$ are generated in degrees less than $\sum_{i=1}^n \ell_i a_i$ and $\delta_i + \sum_{i=1}^n \ell_i a_i$ respectively, Lemma \ref{lem:composite-map} implies that $\ker(\phi)$ is generated in degrees less than
\[
	\max \left(\sum_{i=1}^n \ell_i a_i, \delta_i + \sum_{i=1}^n \ell_i a_i \right).
\]

@@

\begin{thm}\label{P-m-generators/relations}
Let $D = \sum_{i=1}^n \alpha_i C_i$ for curves $C_i$ on $\mathbb{P}^m$.  Let $f_i$ be homogeneous polynomials in $x_0, \ldots, x_m$ such that $C_i = v(f_i)$ for all $i$; suppose $\deg(f_i) = a_i$. 
Then $R_D$ is generated in degree at most 
\[
	\sum_{i=1}^n \ell_i a_i
\]
with relations generated in degree at most
\[
	\max \left(\sum_{i=1}^n \ell_i a_i, \delta_i + \sum_{i=1}^n \ell_i a_i \right).
\]
\end{thm}

\section{Canonical Rings on Hirzebruch surfaces}
Let $D=\sum_{i=1}^m \alpha_i C_i \in \mathbb{Q} \otimes \di(\hirz_
k)$ for some curve $C_i$ of $\hirz_k$, such
that $C_1, \ldots, C_4$ are independent curves with bi-degrees $(1,0)
, (1,0), (0,1)$, and $(0,1)$ respectively. Let $C_i = V(f_i)$ where
$f_i \in \mathscr{O}(\bida_i, \bidb_i)$.  Suppose $f_1$ and $f_2$ are independent linear
polynomials in $x_0$ and $x_1$ and $f_3$ and $f_4$ are independent
linear polynomials in $y_0$ and $y_1$.

\begin{defn}
Define 
\begin{equation}\label{eqn:define-T=(D)}
	\Te(D) = \left\{i \in \{1, \ldots, m\}: \bida_i \sum_{k=1}^m \bidb_k 
\alpha_k = \bidb_i \sum_{k=1}^m \bida_k \alpha_k \right\},
\end{equation}

\begin{equation}\label{eqn:define-T+(D)}
	\Tp(D) = \left\{ i \in \{1, \ldots, m\}:  \vphantom{\sum_{k=1}^m} 
\bida_i \sum_{k=1}^m \alpha_k \bidb_k > \bidb_i \sum_{k=1}^m \alpha_k \bida_k 
\right\},
\end{equation}

\noindent
and
\begin{equation}\label{eqn:define-T-(D)}
	\Tm(D) = \left\{ j \in \{1, \ldots, m\}: \bida_j \sum_{k=1}^m \alpha_
k \bidb_k < \bidb_j \sum_{k=1}^m \alpha_k \bida_k s \right\}.
\end{equation}
\end{defn}
